MachLensFinder - I: An automated Strong Lensing Identification Pipeline

### 2017-01-11 ms.tex
changed focus of motivation to galaxy-galaxy lensing, mass reconstruction,  m/l ratio measurements

% Gravitational lensing as a geometric (?) test of cosmology.  Also
% probes underlying matter distribution in galaxy clusters, sensitive
% to things like recent accretion and environment.
Gravitational lensing occurs when intermediate fluctuations in the
matter density field deflect light from background sources
\citep[see][for a review]{kneibandnatarajan_11}.  In particular, the
deep gravitational potential wells of galaxy clusters provide a strong
lens, with lensed images that can be seen by eye.  Strong
gravitational lensing can manifest as visible giant arcs magnifying
high redshift galaxies \citep{lyndsandpetrosian_86,gladders_etal03},
multiply imaged quasars \citep{walsh_etal79}, and arclets.  Lensing
signatures probe the underlying dark matter distribution of galaxy
clusters \citep{warrenanddye_03}, probe high redshift galaxy formation
\citep{allam_etal07}, and provide a geometric test of cosmology via
comparison of predicted arc abundances and observed abundances
\citep{kochanek_96,chae_03,linder_04}.

% Arc identification can be done visually (e.g. that zooniverse
% thing).  But, in the advent of optical surveys, DES and LSST, we
% will be limited by our ability to automatically process the large
% amounts of data.  Wide field surveys - SPT in millimeter, Herschel
% in submillimeter; high resolution spectroscopic imaging with ALMA;
% optical surveys DES, HSC, LSST, and Euclid (Oguri \& Marshall 2010).
Visual arc identification has been effective through the use of
citizen science platforms (e.g. \citet{marshall_etal16,more_etal16}).
However, in the advent of large surveys, such as the Dark Energy
Survey (DES) and the Large Synoptic Sky Telescope (LSST), we will be
limited by our ability to automatically process large amounts of data
and select likely strong lensing clusters for follow up visual and
spectroscopic confirmation.  The strong lensing sector of
cluster-based cosmology requires an exploration of techniques in image
processing and classification.

% Image identification works: visual, ``robot'' identification -
% ArcFinder, RingFinder..., and citizen science (space warps).  Many
% of these rely on some explicit aspect of the image's morphology
% (give example), filtering, ... ....  Specifics of the methodology
% how the instrument affects the strong lensing signal, and the
% objects of interest.
Over the last decade, infrastructure has emerged for both large scale
visual and automated image classification. By nature, the human eye is
one of the best discriminators for image classification.  {\em
  SpaceWarps} is an example of citizen science based image
classification of strong lensing systems in Canada France Hawaii
Telescope Science (CFHTLS) telescope observations
\citep{marshall_etal16,more_etal16}.  Visual classification has been
quite successful for a dataset like CFHTLS, classifying \change{XX in
  XX time}.  However, future datasets like Euclid and LSST will be
\change{XX} times this size, challenging the scalability of a citizen
science approach.  The next step is to find automated methods with
performance comparable to or better than humans. {\em SpaceWarps} is a
part of the Zooniverse Project (CITE???), which also includes {\em
  Galaxy Zoo}, the citizen science based image classification of
galaxy types \citep{lintott_etal08}.  Galaxy classification is a
successful example of in which machine learning algorithms
successfully train models to classify images with comparable
performance to humans \citep{dieleman_etal15}.

% Describe general robot identification schemes that have been
% implemented
General automated, or ``robot'', identification refers to a classifier
where images are fed into an image processing pipeline that enhances
and extracts characteristic features as parameters on which a pattern
recognition piece is applied.  Examples of this include ArcFinder for
group or cluster scale lens \citep{seidelandbartelmann_07}, Ringfinder
for multiply imaged quasars\citep{gavazzi_etal14}, .  Each pipeline
.... ... and relies on the image morphology where ... ...  Note, that
the method best applied to ... ... and the objects of interest.

@@

We show results of the pipeline on mock galaxy-galaxy lens systems
observed by HST and LSST as respective examples of classifier
performance on optical space- and ground-based observations.
\citet{miraldaescudeandlehar_92} first suggested that massive
ellipticals would be likely frequent strong lensing sources in optical
surveys.  We train and test our pipeline on subsamples from 10,000
mock observed brightest cluster galaxy (BCG) strong lensing systems
and 10,000 BCG non-strong lensing systems.

% Arc finding must optimize for completeness and purity.  
In general, lensed image finding must optimize for completeness and
purity.  In the application of arc abundance comparisons with
cosmological predictions, completeness is important; we must be able
to understand the reasons and frequency of false negatives.  Since
spectroscopic confirmation of background lensed images is expensive,
purity is key; we must minimize false positives that would waste
follow up telescope time.  We discuss the performance of our image
processing and classification pipeline with respect to the
completeness and purity rates of our trained models and illustrate the
scalability of training our models [wc???].

@@

%-------------------------------------------------%
\subsection{Image classification performance}\label{sec:performance}
%-------------------------------------------------%

@@

% TP and TN
Figure ~\ref{fig:HSTclassificationtile} shows a tile of representative
HST-like images classified with the model, HST:N5ppc32cpb4C100.  The
left column show lensing systems, and right column non-lensing
systems.  From top to bottom, the images are ordered from highest
scoring to lowest scoring with the classification model.  The top left
is what we would call a ``true positive'', a lensing system that will
always classified as positive because of its high score.  Most of the
highest scoring lensing systems exhibit similar extended arcs
\todo{corresponding to a large magnification.}  The bottom right is a
``true negative'', a non-lensing system that will likely be classified
as negative because of its low score.  Many of the lowest scoring
non-lensing systems have \todo{fairly small BCGs, which are non likely
  to lens a background galaxy.}

% FN and FP
The bottom left corresponds to a likely ``false negative,'' or a
lensing system that is likely to be mis-classified as a negative.  The
lowest scoring lensing systems tend to have images with smaller
magnification that are minorly distorted, mimicking the shape of
line-of-sight galaxies that the model has learned to ignore.  The top
right corresponds to a likely ``false positive,'' or a non-lensing
system that is likely to be mis-classified as a positive.  False
positives tend to have ``fuzzy'' along the line of sight galaxies that
mimic lensing arcs.

@@

\caption{LSST-like images with a BCG. These illustrate the same image
  classification paradigms as Figure~\ref{fig:HSTclassificationtile}.
  Note, the larger PSF of LSST reduces the resolution of images,
  making the arc-like morphology of faint images less significant and
  increases the appearance of image-BCG blending.}
\label{fig:LSSTBCGclassificationtile}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

@@

\item Images that were difficult for our model to classify typically
  fell in the ..... ....

@@

\acknowledgments CA acknowledges support from the Kavli Institute of
Cosmological Physics, the Enrico Fermi Institute at the University of
Chicago, and the University of Chicago Provost's Office.
\lastpagefootnotes

>>>>

% Gravitational lensing as a geometric (?) test of cosmology.  Also
% probes underlying matter distribution in galaxy clusters, sensitive
% to things like recent accretion and environment.
Gravitational lensing occurs when intermediate fluctuations in the
matter density field deflect light from background sources
\citep[see][for a review]{kneibandnatarajan_11}.  Strong gravitational
lensing can manifest as visible giant arcs magnifying high redshift
galaxies \citep{lyndsandpetrosian_86,gladders_etal03}, multiply imaged
quasars \citep{walsh_etal79}, and arclets.  Lensing signatures probe
the underlying dark matter distribution of the lens
\citep{warrenanddye_03}, high redshift galaxy formation
\citep{allam_etal07}, and provide a geometric test of cosmology via
comparison of predicted arc abundances and observed abundances
\citep{kochanek_96,chae_03,linder_04}.

% Arc identification can be done visually (e.g. that zooniverse
% thing).  But, in the advent of optical surveys, DES and LSST, we
% will be limited by our ability to automatically process the large
% amounts of data.  Wide field surveys - SPT in millimeter, Herschel
% in submillimeter; high resolution spectroscopic imaging with ALMA;
% optical surveys DES, HSC, LSST, and Euclid (Oguri \& Marshall 2010).
Visual arc identification has been effective through the use of
citizen science platforms (e.g. \citet{marshall_etal16,more_etal16}).
However, in the advent of large surveys, such as the Dark Energy
Survey (DES) and the Large Synoptic Sky Telescope (LSST), we will be
limited by our ability to automatically process large amounts of data
and select likely strong lensing candidates for follow up visual and
spectroscopic confirmation.  The strong lensing sector requires an
exploration of techniques in image processing and classification.

% Image identification works: visual, ``robot'' identification -
% ArcFinder, RingFinder..., and citizen science (space warps).  Many
% of these rely on some explicit aspect of the image's morphology
% (give example), filtering, ... ....  Specifics of the methodology
% how the instrument affects the strong lensing signal, and the
% objects of interest.
Over the last decade, infrastructure for both large scale visual and
automated image classification emerged. By nature, the human eye is
one of the best discriminators for image classification.  {\em
  SpaceWarps} is an example of citizen science based image
classification of strong lensing systems in Canada France Hawaii
Telescope Science (CFHTLS) telescope observations
\citep{marshall_etal16,more_etal16}.  Visual classification has been
quite successful for a dataset like CFHTLS, classifying \change{XX in
  XX time}.  However, future datasets like Euclid and LSST will be
\change{XX} times this size, challenging the scalability of a citizen
science approach.  The next step is to find automated methods with
performance comparable to or better than humans. {\em SpaceWarps} is a
part of the Zooniverse Project (CITE???), which also includes {\em
  Galaxy Zoo}, the citizen science based image classification of
galaxy types \citep{lintott_etal08}.  Galaxy classification is a
successful example of in which machine learning algorithms
successfully train models to classify images with comparable
performance to humans \citep{dieleman_etal15}.

% Describe general robot identification schemes that have been
% implemented
General automated, or ``robot'', identification refers to a classifier
where images are fed into an image processing pipeline that enhances
and extracts characteristic features as parameters on which a pattern
recognition piece is applied.  Examples of this include ArcFinder for
group or cluster scale lens \citep{seidelandbartelmann_07}, Ringfinder
for multiply imaged quasars\citep{gavazzi_etal14}, and ???.  Each
pipeline .... ... and relies on the image morphology where ... ...
Note, that the method best applied to ... ... and the objects of
interest.

@@

% Summary of results
We show results of the pipeline on mock galaxy-galaxy lens systems
observed by HST and LSST as respective examples of classifier
performance on optical space- and ground-based observations.
\citet{miraldaescudeandlehar_92} first suggested that massive
ellipticals would be likely frequent strong lensing sources in optical
surveys.  We train and test our pipeline on subsamples from 10,000
mock observed brightest cluster galaxy (BCG) strong lensing systems
and 10,000 BCG non-strong lensing systems.

% Need to optimize for completeness and purity.  Arc counts and mass
% reconstruction.
In general, lensed image finding must optimize for completeness and
purity.  In the application of arc abundance comparisons with
cosmological predictions, completeness is important; we must be able
to understand the reasons and frequency of false negatives.  In the
application of mass reconstruction of the lensing object, completeness
in strong lensing system identifications will allow us to constrain
the mass-to-light ratio for a large number of objects.

% Purity issues
Since spectroscopic confirmation of background lensed images is
expensive, purity is key; we must minimize false positives that would
waste follow up telescope time.  We discuss the performance of our
image processing and classification pipeline with respect to the
completeness and purity rates of our trained models and illustrate the
scalability of training our models [wc???].

@@

%-------------------------------------------------%
\subsection{Image classification performance}\label{sec:performance}
%-------------------------------------------------%
**Redshift makes sense, but only because of the redshift distribution.
--> Zahid+15 quiescent.... this is where redshifts are from.

@@

% TP and TN
Figure ~\ref{fig:HSTclassificationtile} shows a tile of representative
HST-like images classified with the model, HST:N5ppc32cpb4C100.  The
left column show lensing systems, and right column non-lensing
systems.  From top to bottom, the images are ordered from highest
scoring to lowest scoring with the classification model.  The top left
is what we would call a ``true positive'', a lensing system that will
always classified as positive because of its high score.  Most of the
highest scoring lensing systems exhibit similar extended arcs
corresponding to a large magnification (see
Section~\ref{sec:peformance_lensmodel}).  The bottom right is a ``true
negative'', a non-lensing system that will likely be classified as
negative because of its low score.  Many of the lowest scoring
non-lensing systems have \todo{fairly small BCGs, which are non likely
  to lens a background galaxy.}

% FN and FP
The bottom left corresponds to a likely ``false negative,'' or a
lensing system that is likely to be mis-classified as a negative.  The
lowest scoring lensing systems tend to have images with smaller
magnification that are minorly distorted, mimicking the shape of
line-of-sight galaxies that the model has learned to ignore.  The top
right corresponds to a likely ``false positive,'' or a non-lensing
system that is likely to be mis-classified as a positive.  False
positives tend to have ``fuzzy'' along the line of sight galaxies that
mimic lensing arcs.  The false positives should be from bright
galaxies along the line of sight.  \todo{The highest scoring
  non-lensing systems often have fairly bright disk shaped galaxies
  along the line of sight.  Visually, these are also virtually
  indistinguishable from a lensed image and would require
  spectroscopic follow-up.}

@@

\caption{LSST-like images with a BCG. These illustrate the same image
  classification paradigms as Figure~\ref{fig:HSTclassificationtile}.
  Note, the larger PSF of LSST reduces the resolution of images,
  making the arc-like morphology of faint images less significant and
  increases the appearance of image-BCG blending.}
\label{fig:LSSTBCGclassificationtile}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
\subsection{BCG removed}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]
  \centering

@@

%-------------------------------------------------%
\subsubsection{Performance dependence on lens-model parameters}
\label{sec:peformance_lensmodel}
%-------------------------------------------------%

% Summary of subsection
Here, we examine lens-model parameters that affect how well our
pipeline can classify the system.  The lens-model parameters we
examined are the redshift, ellipticity, orientation angle, and
velocity dispersion of the lensing BCG, and also the magnification of
the lensed image compared with its original size.  The magnification
of the lensed image is the most correlated lens-model parameter with
our trained model performance.  The more strongly lensed an image is,
the larger its magnification, and the easier it is for our trained
model to classify.

% Discuss figure
In Figure~\ref{fig:scorevslensparams}, we show the classification
score as a function of image magnification for our HST-like sample.
The left-hand side shows the relation color-coded by the redshift of
the lensing galaxy, and the right-hand side shows the relation
color-coded by the velocity dispersion of the lensing galaxy.  Lensing
systems with images that have magnification $\gsim7$ will almost
certainly be classified as positive with threshold scores above
$\gsim0.5$.  These systems are also those that are most easily
classified by eye.  However, our trained model has varying performance
for systems with lower magnifications.

% Some dependence on redshift and velocity dispersion, which is folded
% into magnitude
Almost \todo{X\%} of lensing systems with magnifications $\lsim7$ will
be falsely identified as non-lenses for classification threshold
scores $\gsim0.5$.  These systems typically have lenses that are at
higher redshifts of $z\gsim0.5$ (lighter colors in left figure) and/or
velocity dispersions lower than $\sigma_v\lsim230.$~km/s (darker
colors in right figure).  Note that lensing galaxies with smaller
velocity dispersions are less massive and therefore have a smaller
efficiency of lensing cross section.  The smaller efficiency means
that a background galaxy is less likely to be strongly lensed with
high magnification.  Also, due to hierarchical structure formation,
galaxies are less massive at higher redshifts, so the trends of model
performance with these three parameters are somewhat degenerate with
one another.

The relationship between our model performance and these lens
parameters indicates that the magnification of lensed images is the
most relevant, and the distribution of image magnification in a data
set will impact trained model performance.  We did not find
correlations in model performance with other lens parameters.  It is
also useful to keep in mind that lensed galaxy images with
magnification $\lsim5$ (e.g. slightly stretched spherical galaxies)
are often visually indistinguishable from elliptical galaxies along
the line of sight.  \todo{We will discuss how differences between the
  lens-model parameter distributions in the training and test sets
  affect performance in a forthcoming paper.}

\mbox{
    \subfigure{\includegraphics[width=\columnwidth,trim=0 0 0 0,
        clip]{figures/HST/or5ppc32cpb4/magnitude_d_colorcodedby_z.pdf}}\hfill
    \subfigure{\includegraphics[width=\columnwidth,trim=0 0 0 0,
        clip]{figures/HST/or5ppc32cpb4/magnitude_d_colorcodedby_velocity_dispersion.pdf}}
} 
\caption{Image classification score as a function of lensed image
  magnification.  Left color-coded by redshift, right is color coded
  by the lens velocity_dispersion.}
\label{fig:scorevslensparams}
\end{figure*}

%-------------------------------------------------%
\subsubsection{Performance sensitivity to line-of-sight galaxies}
\label{sec:peformance_lensmodel}
%-------------------------------------------------%

@@

machine learning -> pipeline, and training set.

Future work with training set tests.

@@

\item Images that were easiest for our model to classify typically
  were lens systems that had high magnification and a BCG with large
  velocity dispersion.

@@

\acknowledgments CA acknowledges support from the Enrico Fermi
Institute at the University of Chicago, and the University of Chicago
Provost's Office. This work was also supported in part by the Kavli
Institute for Cosmological Physics at the University of Chicago
through grant NSF PHY-1125897 and an endowment from the Kavli
Foundation and its founder Fred Kavli.
\lastpagefootnotes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\section{Sample stacked mock LSST images}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\subsection{BCG included}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
